# Executive Summary 

The GOA dusky rockfish is classified as a Tier 3 stock and is assessed using a statistical age-structure model.
This assessment consists of a population model, which uses survey and fishery data to generate a historical time series of population estimates, and a projection model, which uses results from the population model to predict future population estimates and recommended harvest levels. 
The data used in this assessment includes total catch biomass, fishery age and size compositions, trawl survey abundance estimates, and trawl survey age compositions. 
For Gulf of Alaska dusky rockfish in `r year` we present a full assessment with updated assessment and projection model results to recommend harvest levels for the next two years.

## Summary of Changes in Assessment Inputs
Relative to the last full assessment the following substantive changes have been made to assessment inputs: 

 - include survey biomass estimates for `r year-1`, 
 - update survey age compositions with `r year-1` data, 
 - update fishery age compositions with `r year-2` data,  
 - update fishery size compositions with `r year-1` data, and 
 - finale catch values for `r year-2` and `r year-1`, and use preliminary catch for `r year`.

The survey biomass estimate is based upon the Groundfish Assessment Program’s (GAP) Vector Autoregressive Spatio-temporal (VAST) model for the GOA. 
The VAST model input deviates slightly from the GAP default settings in that a lognormal error distribution was implemented instead of the gamma distribution (described in more detail in the *Description of Alternative Models* section below).
In the last few assessments survey biomass from 1984 and 1987 have been included in the survey biomass estimate (though not in the compositional data), however those surveys used different vessels and gear and are not directly comparable to survey data from 1990+. 
Therefore they have been excluded from this assessment going forward.

## Summary of Changes in Assessment Methodology
The following model changes are recommended in the current assessment: extend the age plus group from age-25 to age-30, and extend the length plus group from 47 cm to 52 cm. 

## Summary of Results

A suite of incremental models were run to investigate the effects of removing 1980s survey data from the assessment, and increasing both the length and age plus groups. 

```{r}
data.frame(Model = c("base", 
                     "m15.5a",
                     "m22",
                     "m22.1",
                     "m22.2",
                     "m22.3",
                     "m22a",
                     "m22.1a",
                     "m22.2a",
                     "m22.3a",
                     "m22b",
                     "m22.1b",
                     "m22.2b",
                     "m22.3b"),
           Description = c("2020 model (m15.5a) and results (includes 1980s survey data)",
                           "base model w/data updated through 2022, using GAP default VAST",
                           "m15.5a using GAP default VAST (survey data 1990+)",
                           "m22 w/increased length plus group",
                           "m22 w/increased age plus group",
                           "m22 w/increased age & length plus groups",
                           "m22 using lognormal error VAST (survey data 1990+)",
                           "m22a w/increased length plus group",
                           "m22a w/increased age plus group",
                           "m22a w/increased age & length plus groups",
                           "m22 using design-based survey abundance (1990+)",
                           "m22b w/increased length plus group",
                           "m22b w/increased age plus group",
                           "m22b w/increased age & length plus groups")) %>% 
  flextable::flextable() %>% 
  flextable::padding(padding = 0, part = "all") %>% 
  flextable::autofit() %>% 
  flextable::font(fontname = "Times", part = "all") %>% 
  flextable::fontsize(size = 10)  
```

The author's preferred model is `r pref_mod`, which is the 2020 model with updated data through `r year` and increased age and length plus groups, using a VAST model-based index of survey abundance with lognormal error.
This model generally produces good visual fits to the data and biologically reasonable patterns of recruitment, abundance, and selectivity, and substantially decreases the retrospective Mohn's rho value.

The `r pref_mod` projected age 4+ total biomass for `r year+1` is `r tot` t. 
The recommended ABC for `r year+1` is `r abc` t, the maximum allowable ABC under Tier 3a. 
This ABC is a `r perc`% increase compared to the `r year` ABC of `r abc0` t and a `r perc1`% increase from the projected `r year + 1` ABC from last year. 
However, the past two years the ABC has been on a "stairstep" incremental increase to adjust for the substantially increased abundance when changing to the VAST model for trawl survey biomass in 2020.  
The `r year+1` GOA-wide OFL for dusky rockfish is `r ofl` t. 

The stock is not being subject to overfishing, is not currently overfished, nor is it approaching a condition of being overfished. 

Reference values for dusky rockfish are summarized in the following table:


```{r, child= "01a-main_table.Rmd", eval = T}

```

```{r, child= "01b-apportionment.Rmd", eval = T}

```

```{r, child= "01c-plan_team.Rmd", eval = F}

```


