## Harvest recommendations

### Amendment 56 Reference Points

Amendment 56 to the GOA Groundfish Fishery Management Plan defines the “overfishing level”
(OFL), the fishing mortality rate used to set OFL ($F_{OFL}$), the maximum permissible ABC, and the fishing mortality rate used to set the maximum permissible ABC. 
The fishing mortality rate used to set ABC ($F_{ABC}$) may be less than this maximum permissible level, but not greater. 
Because reliable estimates of reference points related to maximum sustainable yield (MSY) are currently not available but reliable estimates of reference points related to spawning per recruit are available, Northern rockfish in the GOA are managed under Tier 3 of Amendment 56. 
Tier 3 uses the following reference points: $B_{40\%}$, equal to 40% of the equilibrium spawning biomass that would be obtained in the absence of fishing; $F_{35\%}$,,equal to the fishing mortality rate that reduces the equilibrium level of spawning per recruit to 35% of the level that would be obtained in the absence of fishing; and $F_{40\%}$, equal to the fishing mortality rate that reduces the equilibrium level of spawning per recruit to 40% of the level that would be obtained in the absence of fishing.
Estimation of the $B_{40\%}$ reference point requires an assumption regarding the equilibrium level of recruitment. 
In this assessment, it is assumed that the equilibrium level of recruitment is equal to the average of age-2 recruitments between 1979 and `r year - 2`. 
Because of uncertainty in very recent recruitment estimates, we lag 2 years behind model estimates in our projection. 
Other useful biomass reference points which can be calculated using this assumption are $B_{100\%}$ and $B_{35\%}$, defined analogously to $B_{40\%}$. 
The `r year` estimates of these reference points are:

|$B_{100\%}$  |$B_{40\%}$   | $B_{35\%}$  | $F_{40\%}$ | $F_{35\%}$ 
|:----------|:--------------:|:-----------:|:---------:|:---------|
|`r b100`   |   `r b40`   | `r b35`       | `r f40`   | `r f35` |


### Specification of OFL and Maximum Permissible ABC

Female spawning biomass for `r year` is estimated at `r sb1` t. 
This is above the $B_{40\%}$ value of `r b40` t. 
Under Amendment 56, Tier 3, the maximum permissible fishing mortality for ABC is $F_{40\%}$ and fishing mortality for OFL is $F_{35\%}$. 
Applying these fishing mortality rates for `r year`, yields the following ABC and OFL:

|ABC | OFL  | 
|:----------|:--------------:|
|`r abc`   |   `r ofl`   | 

A standard set of projections is required for each stock managed under Tiers 1, 2, or 3 of Amendment 56.
This set of projections encompasses seven harvest scenarios designed to satisfy the requirements of Amendment 56, the National Environmental Policy Act, and the Magnuson-Stevens Fishery Conservation and Management Act (MSFCMA).

For each scenario, the projections begin with the vector of `r year` numbers at age as estimated in the assessment. 
This vector is then projected forward to the beginning of `r year + 1` using the schedules of natural mortality and selectivity described in the assessment and the best available estimate of total (year-end) catch for `r year`. 
In each subsequent year, the fishing mortality rate is prescribed on the basis of the spawning biomass in that year and the respective harvest scenario. 
In each year, recruitment is drawnfrom an inverse Gaussian distribution whose parameters consist of maximum likelihood estimates determined from recruitments estimated in the assessment. 
Spawning biomass is computed in each year based on the time of peak spawning and the maturity and weight schedules described in the assessment. 
Total catch after `r year` is assumed to equal the catch associated with the respective harvest scenario in all years. 
This projection scheme is run 1,000 times to obtain distributions of possible future stock sizes, fishing mortality rates, and catches.

Five of the seven standard scenarios will be used in an Environmental Assessment prepared in conjunction with the final SAFE. 
These five scenarios, which are designed to provide a range of harvest alternatives that are likely to bracket the final TAC for `r year+1`, are as follow (“$max F_{ABC}$” refers to the maximum permissible value of $F_{ABC}$ under Amendment 56):

 - Scenario 1: In all future years, *F* is set equal to $max F_{ABC}$. (Rationale: Historically, TAC has been constrained by ABC, so this scenario provides a likely upper limit on future TACs.)
 
 - Scenario 2: In `r year` and `r year + 1`, *F* is set equal to a constant fraction of $max F_{ABC}$, where this fraction is equal to the ratio of the realized catches in `r year-3`-`r year-1` to the ABC recommended in the assessment for each of those years. 
For the remainder of the future years, maximum permissible ABC is used. (Rationale: In many fisheries the ABC is routinely not fully utilized, so assuming an average ratio catch to ABC will yield more realistic projections.)

 - Scenario 3: In all future years, *F* is set equal to 50% of $max F_{ABC}$. (Rationale: This scenario provides a likely lower bound on $F_{ABC}$ that still allows future harvest rates to be adjusted downward when stocks fall below reference levels.)

 - Scenario 4: In all future years, *F* is set equal to the `r year-5`-`r year-1` average *F*. (Rationale: For some stocks, TAC can be well below ABC, and recent average *F* may provide a better indicator of $F_{TAC}$ than $F_{ABC}$.)

 - Scenario 5: In all future years, *F* is set equal to zero. (Rationale: In extreme cases, TAC may be set at a level close to zero.)

Two other scenarios are needed to satisfy the MSFCMA’s requirement to determine whether a stock is currently in an overfished condition or is approaching an overfished condition. 
These two scenarios are as follows (for Tier 3 stocks, the MSY level is defined as $B_{35\%}$):

 - Scenario 6: In all future years, *F* is set equal to $F_{OFL}$. (Rationale: This scenario determines whether a stock is overfished. If the stock is expected to be 1) above its MSY level in `r year` or 2) above ½ of its MSY level in `r year` and above its MSY level in `r year+10` under this scenario, then the stock is not overfished.)

 - Scenario 7: In `r year+1` and `r year+2`, *F* is set equal to max $F_{ABC}$, and in all subsequent years F is set equal to FOFL. (Rationale: This scenario determines whether a stock is approaching an overfished condition. If the stock is 1) above its MSY level in `r year+2` or 2) above 1/2 of its MSY level in `r year+2` and expected to be above its MSY level in `r year+12` under this scenario, then the stock is not approaching an overfished condition.)
 
Spawning biomass, fishing mortality, and yield are tabulated for the seven standard projection scenarios (Table `r params$chapter`-\@ref(tab:proj)). 
For projections in Scenario 2 (Author’s *F*); we use pre-specified catches to increase accuracy of short-term projections in fisheries where the catch is usually less than the ABC. 
This was suggested to help management with setting preliminary ABCs and OFLs for two-year ahead specifications.

During the 2006 CIE review, it was suggested that projections should account for uncertainty in the entire assessment, not just recruitment from the endpoint of the assessment. 
We continue to present an alternative projection scenario using the uncertainty of the full assessment model harvesting at the same estimated yield ratio (0.67) as Scenario 2, except for all years instead of the next two. 
This projection propagates uncertainty throughout the entire assessment procedure and is based on an MCMC chain of 10 million. 
The projection shows wide credibility intervals on future spawning biomass (Figure 12-\@ref(fig:fswath)). 
The $B_{35\%}$ and $B_{40\%}$ reference points are based on the 1981-`r year-4` age-4 recruitments, and this projection predicts that the median spawning biomass will decrease quickly until average recruitment is attained.


## Risk Table and ABC recommendation


The SSC in its December 2018 minutes recommended that all assessment authors use the risk table when determining whether to recommend an ABC lower than the maximum permissible. 
The following template is used to complete the risk table:

```{r}
library(flextable)
library(tidyverse)

data.frame(level = c("Level 1: Normal",
                     "Level 2: Substantially increased concerns",
                     "Level 3: Major Concern",
                     "Level 4: Extreme concern"),
           assess = c("Typical to moderately increased uncertainty/minor unresolved issues in assessment.", 
                      "Substantially increased assessment uncertainty/ unresolved issues.",
                      "Major problems with the stock assessment; very poor fits to data; high level of uncertainty; strong retrospective bias.",
                      "Severe problems with the stock assessment; severe retrospective bias. Assessment considered unreliable."),
           pop = c("Stock trends are typical for the stock; recent recruitment is within normal range.",
                   "Stock trends are unusual; abundance increasing or decreasing faster than has been seen recently, or recruitment pattern is atypical.",
                   "Stock trends are highly unusual; very rapid changes in stock abundance, or highly atypical recruitment patterns.",
                   "Stock trends are unprecedented; More rapid changes in stock abundance than have ever been seen previously, or a very long stretch of poor recruitment compared to previous patterns."),
           enviro = c("No apparent environmental/ecosystem concerns",
                      "Some indicators showing adverse signals relevant to the stock but the pattern is not consistent across all indicators.",
                      "Multiple indicators showing consistent adverse signals a) across the same trophic level as the stock, and/or b) up or down trophic levels (i.e., predators and prey of the stock)",
                      "Extreme anomalies in multiple ecosystem indicators that are highly likely to impact the stock; Potential for cascading effects on other ecosystem components"),
           fish = c("No apparent fishery/resource-use performance and/or behavior concerns",
                    "Some indicators showing adverse signals but the pattern is not consistent across all indicators",
                    "Multiple indicators showing consistent adverse signals a) across different sectors, and/or b) different gear types",
                    "Extreme anomalies in multiple performance  indicators that are highly likely to impact the stock")) %>% 
  flextable() %>% 
    font(fontname = "Times", part = "all") %>% 
    width(j = ~ assess, width = 1.5) %>% 
    width(j = ~ pop, width = 2) %>% 
    width(j = ~ enviro, width = 1.5) %>% 
    width(j = ~ fish, width = 1.5) %>% 
    set_header_labels(level = "",
                    assess = "Assessment-related considerations",
                    pop = "Population dynamics considerations",
                    enviro = "Environmental/ecosystem considerations",
                    fish = "Fishery Performance") %>% 
  fontsize(size = 10) %>% 
  theme_vanilla() %>% 
  italic(part = "header") 
```  
The table is applied by evaluating the severity of four types of considerations that could be used to support a scientific recommendation to reduce the ABC from the maximum permissible. 
These considerations are stock assessment considerations, population dynamics considerations, environmental/ecosystem considerations, and fishery performance. 
Examples of the types of concerns that might be relevant include the following: 

#### Assessment considerations

Level 2. The GOA trawl survey was conducted in `r year-1` as expected, and fishery and survey age and length composition data have been incorporated with the expected range of data made available on time for incorporation into the 2020 stock assessment. 
The assessment model produces reasonable fits to the survey abundance index and compositional data, but the model results are sensitive to the VAST biomass index and the additional year of survey data. 
The low variance of the VAST model configuration compared to uncertainty from the design-based methods, coupled with high `r year-1` survey biomass estimate has resulted in continued high levels of total and spawning biomass. 
In addition, all VAST configurations examined result in increased biomass and ABC estimates relative to design-based abundance inputs. 
It is worth noting that catch has not similarly increased.
Because of the sensitivity of the assessment model to the scale of the VAST model biomass estimates, as well as the low uncertainty from this method, we recommend a level 2 concern for assessment considerations.  


#### Population dynamics considerations

Level 2. There is a continued increase in estimated survey biomass for dusky rockfish when using VAST inputs of abundance in the assessment. 
Fishery and survey age compositions for the most recent year are both relatively uniformly distributed across ages with some increased proportions of 8-11 year old fish over the previous compositions. 
The assessment model estimates a high age-4 recruitment for 2010 and these recruits would be 13 years old in 2019. 
This lines up somewhat imperfectly with both the observed age compositions for the fishery and survey which show a higher proportion age 11 fish in 2018 (fishery) and age 11 and 12 fish in 2019 (survey). 
Rockfish aging is challenging, and some smearing across ages is expected. 
Skip spawning has been observed for this species [@Conrath2019], though the spatial and temporal extent is unknown.
However, preliminary investigations that incorporate skip spawning in maturity estimates lead to a reduction is spawning biomass and associated ABC.  
For these reasons we have given this risk table factor a level 2 concern for population dynamics considerations though make no recommendation for a reduction in ABC.
 

#### Environmental/Ecosystem considerations

Level 1. Environmental mechanisms for changes in survival and productivity of dusky rockfish remain unknown, though changes in water temperature and currents could have effects on prey abundance and success of transition of rockfish from pelagic to demersal stage. 
Predator effects would likely be more important on larval, post-larval, and small juvenile slope rockfish, but there is insufficient information on these life stages and their predators to inform a conclusion. 
Additionally, changes in bottom habitat due to natural or anthropogenic causes could alter survival rates by altering available shelter, prey, or other functions. 
Estimates of structural epifauna habitat (with non-targeted data) have recently been in decline.
However, given the continued lack of biological and habitat information for dusky rockfish, we scored this category as Level 1, as the level of concern has not changed.

#### Fishery performance

Level 1. Catches are well below ABC for 2022, which matches the historical trend of the fishery catch rarely approaching ABC (Table 12-\@ref(tab:catch)). 
Dusky rockfish are caught with a number of other rockfish species, so TAC levels for Pacific Ocean perch and northern rockfish, as well as prohibited species catch restrictions (i.e. salmon) can also affect fishery realization of the full TAC. 
In addition, dusky catches can be influenced by the price, and current prices are relatively low (J. Bonney, pers. Comm. Oct 2020). 
For these reasons we have given this risk table factor a ‘level 1, normal’ rank and do not suggest there is reason to suggest a reduction in ABC based on fishery performance considerations.


#### Summary and ABC recommendation


```{r}
l1 = "Level 1: No increased concerns"
l2 = "Level 2: Substantially increased concerns"
l3 = "Level 3: Major concern"
l4 = "Level 4: Extreme concern"


data.frame(assess = l2,
           pop = l2,
           enviro = l1,
           fish = l1) %>% 
  flextable() %>% 
    width(j = ~ assess, width = 1.5) %>% 
    width(j = ~ pop, width = 1.5) %>% 
    width(j = ~ enviro, width = 1.5) %>% 
    width(j = ~ fish, width = 1.5) %>% 
    set_header_labels(level = "",
                    assess = "Assessment-related considerations",
                    pop = "Population dynamics considerations",
                    enviro = "Environmental/ecosystem considerations",
                    fish = "Fishery Performance") %>% 
  fontsize(size = 10) %>% 
  font(fontname = "Times", part = "all") %>% 
  italic(part = "header") 
```
The GOA dusky rockfish assessment appears to fit available data well, the 2021 GOA trawl survey was undertaken as planned and data are included in this year’s assessment, and the fishery and environmental considerations appear to be within normal bounds. 
We have some concerns about the estimated increase in biomass and resulting increase in ABC. 
The VAST-based abundance index has low uncertainty which may be driving the estimated increase in biomass and ABC. 
Additionally, there are unknown levels of skip spawning withing this population, the implications of which are not fully understood, though any increase in skip spawning reduces the spawning population, and therefore ABC estimate. 
Because GOA dusky rockfish ABC is not historically fully utilized and because there is some evidence of recruitment from age compositions, we are not recommending a reduction in ABC at this time. 
We anticipate that we will monitor the survey abundance estimates, catch rates, and explore skip spawning more fully in the next assessment.

### Area Allocation of Harvests

The random effects model was fit to the survey design-based biomass estimates (with associated variance) for the Western, Central, and Eastern GOA. 
The random effects model estimates a process error parameter (constraining the variability of the modeled estimates among years) and random effects parameters in each year modeled. 
The fit of the random effects model to survey biomass in each area is shown in (Figure 12-\@ref(fig:fre)). 
In general the random effects model fits the area-specific survey biomass reasonably well. 
Using the random effects model estimates of survey biomass, the apportionment results in 1.9% for the Western area, 96.6% for the Central area, and 1.5% for the Eastern area. 
The changes in apportionment in `r year` compared to `r year-2` can be attributed to a concentration in biomass from the bottom trawl survey biomass to the Central area. 
This results in recommended ABC’s of 149 t for the Western area, 7,647 t for the Central area, and 121 t for the Eastern area.

Because the Eastern area is divided into two management areas for dusky rockfish, i.e., the West Yakutat area (area between 147 degrees W. longitude and 140 degrees W. longitude) and the East Yakutat/Southeast Outside area (area east of 140 degrees W. longitude), the ABC for this management group in the Eastern area must be further apportioned between these two smaller areas. 
In an effort to balance uncertainty with associated costs to the fishing industry, the GOA Plan Team has recommended that apportionment to the two smaller areas in the Eastern GOA be based on the upper 95% confidence limit of the weighted average of the estimates of the Eastern GOA biomass proportion that is in the West Yakutat area. 
The upper 95% confidence interval of this proportion is 0.55, so that the dusky rockfish ABC for West Yakutat is 149 t, and the ABC for East Yakutat/Southeast Outside is 121 t.
Based on the definitions for overfishing in Amendment 44 in Tier 3a (i.e., $F_{OFL} = F_{35\%}=0.11$), the `r year+1` overfishing (OFL) is set equal to 9,638 t for dusky rockfish in the GOA.


### Status Determination

Under the MSFCMA, the Secretary of Commerce is required to report on the status of each U.S. fishery with respect to overfishing. 
This report involves the answers to three questions: 1) Is the stock being subjected to overfishing? 2) Is the stock currently overfished? 3) Is the stock approaching an overfished condition?

*Is the stock being subjected to overfishing?* The official catch estimate for the most recent complete year (`r year-1`) is `r c0`  t. 
This is less than the `r year-1` OFL of `r ofl0` t. 
Therefore, the stock is not being subjected to overfishing.

Harvest Scenarios #6 and #7 are intended to permit determination of the status of a stock with respect to its minimum stock size threshold (MSST). 
Any stock that is below its MSST is defined to be overfished. 
Any stock that is expected to fall below its MSST in the next two years is defined to be approaching an overfished condition. 
Harvest Scenarios #6 and #7 are used in these determinations as follows:

*Is the stock currently overfished?* This depends on the stock’s estimated spawning biomass in `r year`:

 a. If spawning biomass for `r year` is estimated to be below ½ $B_{35\%}$, the stock is below its MSST.
 b. If spawning biomass for `r year` is estimated to be above $B_{35\%}$ the stock is above its MSST.
 c. If spawning biomass for `r year` is estimated to be above ½ $B_{35\%}$ but below $B_{35\%}$, the stock’s status relative to MSST is determined by referring to harvest Scenario #6 (Table 12-\@ref(tab:proj)). 
If the mean spawning biomass for `r year+12` is below $B_{35\%}$, the stock is below its MSST. 
Otherwise, the stock is above its MSST.

*Is the stock approaching an overfished condition?* This is determined by referring to harvest Scenario #7:

 a. If the mean spawning biomass for `r year+2` is below 1/2 $B_{35\%}$, the stock is approaching an overfished condition.
 b. If the mean spawning biomass for `r year+2` is above $B_{35\%}$, the stock is not approaching an overfished condition.
 c. If the mean spawning biomass for `r year+2` is above 1/2 $B_{35\%}$ but below $B_{35\%}$, the determination depends on the mean spawning biomass for `r year+12` 
If the mean spawning biomass for `r year+12` is below $B_{35\%}$, the stock is approaching an overfished condition. 
Otherwise, the stock is not approaching an overfished condition.

Based on the above criteria and projections in Table 12-\@ref(tab:proj), the stock is not overfished and is not approaching an overfished condition. 
The fishing mortality that would have produced a catch for last year equal to last year’s OFL is `r best_f`.


