## Harvest recommendations

(Items in this section pertain to the authors’ recommended model or approach. 
If the structure of the recommended model or approach differs substantively from the model or approach most recently accepted by the SSC after reviewing either last year’s final SAFE report or the current year’s preliminary SAFE report, a set of parallel results for the previously accepted model or approach should be included in an attachment.)


### Amendment 56 Reference Points

Amendment 56 to the GOA Groundfish Fishery Management Plan defines the “overfishing level”
(OFL), the fishing mortality rate used to set OFL ($F_{OFL}$), the maximum permissible ABC, and the fishing mortality rate used to set the maximum permissible ABC. 
The fishing mortality rate used to set ABC ($F_{ABC}$) may be less than this maximum permissible level, but not greater. 
Because reliable estimates of reference points related to maximum sustainable yield (MSY) are currently not available but reliable estimates of reference points related to spawning per recruit are available, Northern rockfish in the GOA are managed under Tier 3 of Amendment 56. 
Tier 3 uses the following reference points: $B_{40\%}$, equal to 40% of the equilibrium spawning biomass that would be obtained in the absence of fishing; $F_{35\%}$,,equal to the fishing mortality rate that reduces the equilibrium level of spawning per recruit to 35% of the level that would be obtained in the absence of fishing; and $F_{40\%}$, equal to the fishing mortality rate that reduces the equilibrium level of spawning per recruit to 40% of the level that would be obtained in the absence of fishing.
Estimation of the $B_{40\%}$ reference point requires an assumption regarding the equilibrium level of recruitment. 
In this assessment, it is assumed that the equilibrium level of recruitment is equal to the average of age-2 recruitments between 1979 and `r year - 2`. 
Because of uncertainty in very recent recruitment estimates, we lag 2 years behind model estimates in our projection. 
Other useful biomass reference points which can be calculated using this assumption are $B_{100\%}$ and $B_{35\%}$, defined analogously to $B_{40\%}$. 
The `r year` estimates of these reference points are:

**TABLE HERE**


### Specification of OFL and Maximum Permissible ABC

Female spawning biomass for `r year` is estimated at **40,462** t. 
This is above the $B_{40\%}$ value of **33,933** t. 
Under Amendment 56, Tier 3, the maximum permissible fishing mortality for ABC is $F_{40\%}$ and fishing mortality for OFL is $F_{35\%}$. 
Applying these fishing mortality rates for `r year`, yields the following ABC and OFL:

**TABLE HERE**


A standard set of projections is required for each stock managed under Tiers 1, 2, or 3 of Amendment 56.
This set of projections encompasses seven harvest scenarios designed to satisfy the requirements of Amendment 56, the National Environmental Policy Act, and the Magnuson-Stevens Fishery Conservation and Management Act (MSFCMA).

For each scenario, the projections begin with the vector of `r year` numbers at age as estimated in the assessment. 
This vector is then projected forward to the beginning of `r year + 1` using the schedules of natural mortality and selectivity described in the assessment and the best available estimate of total (year-end) catch for `r year`. 
In each subsequent year, the fishing mortality rate is prescribed on the basis of the spawning biomass in that year and the respective harvest scenario. 
In each year, recruitment is drawnfrom an inverse Gaussian distribution whose parameters consist of maximum likelihood estimates determined from recruitments estimated in the assessment. 
Spawning biomass is computed in each year based on the time of peak spawning and the maturity and weight schedules described in the assessment. 
Total catch after `r year` is assumed to equal the catch associated with the respective harvest scenario in all years. 
This projection scheme is run 1,000 times to obtain distributions of possible future stock sizes, fishing mortality rates, and catches.

Five of the seven standard scenarios will be used in an Environmental Assessment prepared in conjunction with the final SAFE. 
These five scenarios, which are designed to provide a range of harvest alternatives that are likely to bracket the final TAC for 2019, are as follow (“$max F_{ABC}$” refers to the maximum permissible value of $F_{ABC}$ under Amendment 56):

 - Scenario 1: In all future years, *F* is set equal to $max F_{ABC}$. (Rationale: Historically, TAC has been constrained by ABC, so this scenario provides a likely upper limit on future TACs.)
 
 - Scenario 2: In `r year` and `r year + 1`, *F* is set equal to a constant fraction of $max F_{ABC}$, where this fraction is equal to the ratio of the realized catches in `r year-3`-`r year-1` to the ABC recommended in the assessment for each of those years. 
For the remainder of the future years, maximum permissible ABC is used. (Rationale: In many fisheries the ABC is routinely not fully utilized, so assuming an average ratio catch to ABC will yield more realistic projections.)

 - Scenario 3: In all future years, *F* is set equal to 50% of $max F_{ABC}$. (Rationale: This scenario provides a likely lower bound on FABC that still allows future harvest rates to be adjusted downward when stocks fall below reference levels.)

 - Scenario 4: In all future years, *F* is set equal to the 2013-2017 average *F*. (Rationale: For some stocks, TAC can be well below ABC, and recent average *F* may provide a better indicator of $F_{TAC}$ than $F_{ABC}$.)

 - Scenario 5: In all future years, *F* is set equal to zero. (Rationale: In extreme cases, TAC may be set at a level close to zero.)

Two other scenarios are needed to satisfy the MSFCMA’s requirement to determine whether a stock is currently in an overfished condition or is approaching an overfished condition. 
These two scenarios are as follows (for Tier 3 stocks, the MSY level is defined as $B_{35\%}$):

 - Scenario 6: In all future years, *F* is set equal to $F_{OFL}$. (Rationale: This scenario determines whether a stock is overfished. If the stock is expected to be 1) above its MSY level in 2018 or 2) above ½ of its MSY level in 2018 and above its MSY level in 2028 under this scenario, then the stock is not overfished.)

 - Scenario 7: In `r year` and `r year + 1`, *F* is set equal to max $F_{ABC}$, and in all subsequent years F is set equal to FOFL. (Rationale: This scenario determines whether a stock is approaching an overfished condition. If the stock is 1) above its MSY level in 2020 or 2) above 1/2 of its MSY level in 2020 and expected to be above its MSY level in 2030 under this scenario, then the stock is not approaching an overfished condition.)
 
Spawning biomass, fishing mortality, and yield are tabulated for the seven standard projection scenarios **ref (Table 10.16)**. 
The difference for this assessment for projections is in Scenario 2 (Author’s *F*); we use pre-specified catches to increase accuracy of short-term projections in fisheries where the catch is usually less than the ABC. 
This was suggested to help management with setting preliminary ABCs and OFLs for two-year ahead specifications.

During the 2006 CIE review, it was suggested that projections should account for uncertainty in the entire assessment, not just recruitment from the endpoint of the assessment. 
We continue to present an alternative projection scenario using the uncertainty of the full assessment model harvesting at the same estimated yield ratio (0.67) as Scenario 2, except for all years instead of the next two. 
This projection propagates uncertainty throughout the entire assessment procedure and is based on an MCMC chain of 1,000,000. 
The projection shows wide credibility intervals on future spawning biomass **ref (Figure 12-16)**. 
The $B_{35%}$ and $B_{40%}$ reference points are based on the 1981-`r year-4` age-4 recruitments, and this projection predicts that the median spawning biomass will decrease quickly until average recruitment is attained.


## Risk Table amnd ABC recommendation


The SSC in its December 2018 minutes recommended that all assessment authors use the risk table when determining whether to recommend an ABC lower than the maximum permissible. The following template is used to complete the risk table:

```{r}
library(flextable)
library(tidyverse)

data.frame(level = c("Level 1: Normal",
                     "Level 2: Substantially increased concerns",
                     "Level 3: Major Concern",
                     "Level 4: Extreme concern"),
           assess = c("Typical to moderately increased uncertainty/minor unresolved issues in assessment.", 
                      "Substantially increased assessment uncertainty/ unresolved issues.",
                      "Major problems with the stock assessment; very poor fits to data; high level of uncertainty; strong retrospective bias.",
                      "Severe problems with the stock assessment; severe retrospective bias. Assessment considered unreliable."),
           pop = c("Stock trends are typical for the stock; recent recruitment is within normal range.",
                   "Stock trends are unusual; abundance increasing or decreasing faster than has been seen recently, or recruitment pattern is atypical.",
                   "Stock trends are highly unusual; very rapid changes in stock abundance, or highly atypical recruitment patterns.",
                   "Stock trends are unprecedented; More rapid changes in stock abundance than have ever been seen previously, or a very long stretch of poor recruitment compared to previous patterns."),
           enviro = c("No apparent environmental/ecosystem concerns",
                      "Some indicators showing adverse signals relevant to the stock but the pattern is not consistent across all indicators.",
                      "Multiple indicators showing consistent adverse signals a) across the same trophic level as the stock, and/or b) up or down trophic levels (i.e., predators and prey of the stock)",
                      "Extreme anomalies in multiple ecosystem indicators that are highly likely to impact the stock; Potential for cascading effects on other ecosystem components"),
           fish = c("No apparent fishery/resource-use performance and/or behavior concerns",
                    "Some indicators showing adverse signals but the pattern is not consistent across all indicators",
                    "Multiple indicators showing consistent adverse signals a) across different sectors, and/or b) different gear types",
                    "Extreme anomalies in multiple performance  indicators that are highly likely to impact the stock")) %>% 
  flextable() %>% 
    font(fontname = "Times", part = "all") %>% 
    width(j = ~ assess, width = 1.5) %>% 
    width(j = ~ pop, width = 2) %>% 
    width(j = ~ enviro, width = 1.5) %>% 
    width(j = ~ fish, width = 1.5) %>% 
    set_header_labels(level = "",
                    assess = "Assessment-related considerations",
                    pop = "Population dynamics considerations",
                    enviro = "Environmental/ecosystem considerations",
                    fish = "Fishery Performance") %>% 
  font(fontname = "Times", part = "all") %>% 
  fontsize(size = 10) %>% 
  theme_vanilla() %>% 
  italic(part = "header") 
```  
The table is applied by evaluating the severity of four types of considerations that could be used to support a scientific recommendation to reduce the ABC from the maximum permissible. 
These considerations are stock assessment considerations, population dynamics considerations, environmental/ecosystem considerations, and fishery performance. Examples of the types of concerns that might be relevant include the following: 

  1. “Assessment considerations—data-inputs: biased ages, skipped surveys, lack of fishery-independent trend data; model fits: poor fits to fits to fishery or survey data, inability to simultaneously fit multiple data inputs; model performance: poor model convergence, multiple minima in the likelihood surface, parameters hitting bounds; estimation uncertainty: poorly-estimated but influential year classes; retrospective bias in biomass estimates.
  2. “Population dynamics considerations—decreasing biomass trend, poor recent recruitment, inability of the stock to rebuild, abrupt increase or decrease in stock abundance.
  3. “Environmental/ecosystem considerations—adverse trends in environmental/ecosystem indicators, ecosystem model results, decreases in ecosystem productivity, decreases in prey abundance or availability, increases or increases in predator abundance or productivity.
  4. “Fishery performance—fishery CPUE is showing a contrasting pattern from the stock biomass trend, unusual spatial pattern of fishing, changes in the percent of TAC taken, changes in the duration of fishery openings.”



#### Assessment considerations

Level 2. The GOA trawl survey was conducted in 2019 as expected, and fishery and survey age and length composition data have been incorporated with the expected range of data made available on time for incorporation into the 2020 stock assessment. 
The assessment model produces reasonable fits to the survey abundance index and compositional data, but the model results are sensitive to the geospatial model biomass index and the additional year of survey data. 
The low variance of the geospatial model configuration, coupled with high 2019 survey biomass estimate has resulted in a large increase in total and spawning biomass. 
There is also a strong positive retrospective bias towards decreasing spawning biomass in recent years **(Mohn’s rho = 0.51)**. 
In addition, all geospatial index configurations examined result in increased biomass and ABC estimates. 
The large increase in biomass, and subsequently ABC, that are primarily driven by the attributes of the geospatial model (low estimated variance in biomass) in conjunction with a strong retrospective pattern that decreases biomass through time provide conflicting trends in this assessment. 
For these reasons we have given this risk table factor a ‘level 2, substantially increase concerns’ rank.

#### Population dynamics considerations

Level 1. There is a large increase in the 2019 survey biomass estimate for dusky rockfish in both the design-based and geospatial indices. 
Fishery and survey age compositions for the most recent year are both relatively uniformly distributed across ages with some increased proportions of 8-11 year old fish over the previous compositions. 
The assessment model estimates a high age-4 recruitment for 2010 and these recruits would be 13 years old in 2019. 
This lines up somewhat imperfectly with both the observed age compositions for the fishery and survey which show a higher proportion age 11 fish in 2018 (fishery) and age 11 and 12 fish in 2019 (survey). 
Rockfish aging is challenging, and some smearing across ages is expected. 
For these reasons we have given this risk table factor a ‘level 1, normal’ rank and do not suggest there is reason to suggest a reduction in ABC based on population dynamics considerations.

#### Environmental/Ecosystem considerations

Level 1. Dusky rockfish are benthic dwellers as adults, with a pelagic then juvenile stage in the Gulf of Alaska (GOA) (Love et al. 2002).
The limited information available on temperature, zooplankton, and conditions of other marine species indicate average foraging and growing conditions for the zooplanktivorous juvenile rockfish during 2020. Heat wave conditions occurred during 2020 but were not as severe as 2019 during the summer and fall in the GOA [@Watson2020]. 
Sea surface temperatures were about 1°C above normal in the western GOA and average in the eastern GOA during the 2020 summer (Alaska Center for Climate Assessment & Policy ACCAP, Thoman personal communication). 
Inside waters of the GOA were slightly more anomalously warm than offshore temperatures (ACCAP). Offshore of Seward, waters above the continental shelf at GAK1 on the Seward line remained anomalously warm (0.5°C) at 200-250 m depth in 2020 but cooler than 2019 [@Danielsen2020]. 
Along the GOA slope, the AFSC Longline Survey Subsurface Temperature Index indicates above average temperatures at the surface and at depth (250 m) in 2020 relative to the 2005-2019 time series and cooler temperatures in 2020 relative to 2019 (Siwicke personal communication). 
In the inside waters, Prince William Sound has remained warm since 2014 [@Campbell2020]. 
However, for the inside waters of the eastern GOA, the top 20 m temperatures of Icy Strait in northern southeast Alaska during summer were slightly below average (8.8°C) in 2020 relative to the 23 year time series (1997-2019) [@Fergusson2020]. 
It is reasonable to expect that the recent heat wave conditions and current return to cooler temperatures would not adversely impact age-0 rockfish in pelagic waters during a time when they are growing to a size that may promote over winter survival, however, it is unknown what this impact will be. 
Further, a recent study published on the U.S. West Coast suggests that the warming that occurred during 2014-2016 may have been beneficial for rockfish recruitment [@Morgan2019].

The known primary prey of the adult dusky rockfish include euphausiids and Pacific sand lance in the GOA [@Byerly2001; @Yang1993; @Yang2000]. 
Warm conditions tend to be associated with zooplankton communities that are dominated by smaller and less lipid rich species in the GOA [@Kimmel2019].
There was limited information on zooplankton in 2020. 
In Icy Strait, northern southeast Alaska, the lipid content of all zooplankton taxa combined examined during 2020 was average for the time series (1997-2020) and similar to 2019. 
By taxa, lipid content was above average for the large calanoid copepods, average for hyperiid amphipods, but lower than average for euphausiids, small copepods and gastropods indicating average the nutritional quality of the prey field utilized by larval and juvenile fishes in the nearshore habitats [@Fergusson2020]. 
In the western GOA, the mean biomass of large calanoids and euphausiids averaged over the top 100m south of Seward Alaska during May were about average in 2020 relative to the time series, 1998-2019 [@Hopcroft2020]. 
On the outer edge of the continental shelf in the central Gulf of Alaska, the breeding success of piscivorous seabirds on Middleton Island as an indication for foraging success and nutrient-rich prey was above average indicating good ocean conditions during 2020 [@Hatch2020]. 
Sand lance was observed in low proportions in the diets of surface feeding birds and moderate proportions in the diets of diving sea birds, similar proportions to 2019 [@Hatch2020]. 
Mean abundance of larval sand lance in spring 2019 was elevated to levels last seen around 2007, indicating potentially increased abundance of age-1 sand lance in 2020 (although no data were collected in 2020 [@Rogers2020]. 
Little is known about the impacts of predators and competitors, such as fish and marine mammals, on dusky rockfish. 
However, survival of larvae are thought more related to the abundance and timing of prey availability than predation, due to the lack of rockfish as a prey item [@Love2002; @Yang2003]. 
The 2020 foraging conditions were likely average, although data limited, for the largely zooplanktivorous and piscivorous dusky rockfish in the GOA. Given cooler conditions in 2020 than in 2019 and average densities and body condition of zooplankton with limited information on rockfish, we scored this category as level 1, as normal concern and do not suggest there is reason to suggest a reduction in ABC based on environmental/ecosystem considerations.

#### Fishery performance

Level 1. Catches are well below ABC for 2020, which matches the historical trend of the fishery catch rarely approaching ABC **ref (Table 12.2)**. 
Dusky rockfish are caught with a number of other rockfish species, so TAC levels for Pacific Ocean perch and northern rockfish, as well as prohibited species catch restrictions (i.e. salmon) can also affect fishery realization of the full TAC. 
In addition, dusky catches can be influenced by the price, and current prices are relatively low (J. Bonney, pers. Comm. Oct 2020). 
For these reasons we have given this risk table factor a ‘level 1, normal’ rank and do not suggest there is reason to suggest a reduction in ABC based on fishery performance considerations.


#### Summary and ABC recommendation


```{r}
l1 = "Level 1: No increased concerns"
l2 = "Level 2: Substantially increased concerns"
l3 = "Level 3: Major concern"
l4 = "Level 4: Extreme concern"


data.frame(assess = l1,
           pop = l1,
           enviro = l1,
           fish = l1) %>% 
  flextable() %>% 
    width(j = ~ assess, width = 1.5) %>% 
    width(j = ~ pop, width = 1.5) %>% 
    width(j = ~ enviro, width = 1.5) %>% 
    width(j = ~ fish, width = 1.5) %>% 
    set_header_labels(level = "",
                    assess = "Assessment-related considerations",
                    pop = "Population dynamics considerations",
                    enviro = "Environmental/ecosystem considerations",
                    fish = "Fishery Performance") %>% 
  fontsize(size = 10) %>% 
  font(fontname = "Times", part = "all") %>% 
  italic(part = "header") 
```
We have ranked three categories as ‘Level 1: No apparent concern’ and one as a ‘Level 2, substantially increased concerns’. 
The GOA dusky rockfish assessment appears to fit available data well, the 2021 GOA trawl survey was undertaken as planned and data are included in this year’s assessment, and the fishery and environmental considerations appear to be within normal bounds. 
We have some concerns about the estimated increase in biomass and resulting increase in ABC and the model retrospective pattern. 
The geospatial model-based abundance index has low uncertainty which may be driving the estimated increase in biomass and ABC. 
Because GOA dusky rockfish ABC is not historically fully utilized and because there is an increase in 2021 survey biomass coupled with some evidence of recruitment from age compositions, we are not recommending a reduction in ABC at this time. 
We anticipate that we will monitor the survey abundance estimates, catch rates, and retrospective trends closely for the next assessment.

### Area Allocation of Harvests

The random effects model was fit to the survey design-based biomass estimates (with associated variance) for the Western, Central, and Eastern GOA. 
The random effects model estimates a process error parameter (constraining the variability of the modeled estimates among years) and random effects parameters in each year modeled. 
The fit of the random effects model to survey biomass in each area is shown in **ref figure**. 
In general the random effects model fits the area-specific survey biomass reasonably well. 
**Using the random effects model estimates of survey biomass, the apportionment results in 5.0% for the Western area (down from 21.1% in 2018), 84.4% for the Central area (up from 74.7% in 2018), and 10.6% for the Eastern area (up from 4.2% in 2018). 
The changes in apportionment in `r year` compared to `r year-2` can be attributed to an increase in biomass from the bottom trawl survey biomass in the Central and Eastern areas. 
This results in recommended ABC’s of 355 t for the Western area, 5,993 t for the Central area, and 753 t for the Eastern area.**

Because the Eastern area is divided into two management areas for dusky rockfish, i.e., the West Yakutat area (area between 147 degrees W. longitude and 140 degrees W. longitude) and the East Yakutat/Southeast Outside area (area east of 140 degrees W. longitude), the ABC for this management group in the Eastern area must be further apportioned between these two smaller areas. 
In an effort to balance uncertainty with associated costs to the fishing industry, the GOA Plan Team has recommended that apportionment to the two smaller areas in the Eastern GOA be based on the upper 95% confidence limit of the weighted average of the estimates of the Eastern GOA biomass proportion that is in the West Yakutat area. 
The upper 95% confidence interval of this proportion is **0.82 (up from 0.61 in `r year-2`)**, so that the dusky rockfish ABC for West Yakutat is **617 t**, and the ABC for East Yakutat/Southeast Outside is **136 t ref(Table 12-17)**.
Based on the definitions for overfishing in Amendment 44 in Tier 3a (i.e., $F_{OFL} = F_{35%}=0.114$), the `r year+1` overfishing (OFL) is set equal to **8,655 t** for dusky rockfish in the GOA **ref (Table 12-17)**.


### Status Determination

Under the MSFCMA, the Secretary of Commerce is required to report on the status of each U.S. fishery with respect to overfishing. 
This report involves the answers to three questions: 1) Is the stock being subjected to overfishing? 2) Is the stock currently overfished? 3) Is the stock approaching an overfished condition?

*Is the stock being subjected to overfishing?* The official catch estimate for the most recent complete year (`r year-1`) is *correct this later* r catch %>% filter(Year==year-1) %>% pull(Catch) %>% format(., big.mark = ",")` t. 
This is less than the `r year-1` OFL of 5,402 t. 
Therefore, the stock is not being subjected to overfishing.

Harvest Scenarios #6 and #7 are intended to permit determination of the status of a stock with respect to its minimum stock size threshold (MSST). 
Any stock that is below its MSST is defined to be overfished. 
Any stock that is expected to fall below its MSST in the next two years is defined to be approaching an overfished condition. 
Harvest Scenarios #6 and #7 are used in these determinations as follows:

*Is the stock currently overfished?* This depends on the stock’s estimated spawning biomass in `r year`:

 - a. If spawning biomass for `r year` is estimated to be below ½ $B_{35\%}$, the stock is below its MSST.
 - b. If spawning biomass for `r year` is estimated to be above $B_{35\%}$ the stock is above its MSST.
 - c. If spawning biomass for `r year` is estimated to be above ½ $B_{35\%}$ but below $B_{35\%}$, the stock’s status relative to MSST is determined by referring to harvest Scenario #6 (Table 10.16). 
If the mean spawning biomass for 2028 is below $B_{35\%}$, the stock is below its MSST. 
Otherwise, the stock is above its MSST.

*Is the stock approaching an overfished condition?* This is determined by referring to harvest Scenario #7:

 - a. If the mean spawning biomass for `r year+2` is below 1/2 $B_{35\%}$, the stock is approaching an overfished condition.
 - b. If the mean spawning biomass for `r year+2` is above $B_{35\%}$, the stock is not approaching an overfished condition.
 - c. If the mean spawning biomass for `r year+2` is above 1/2 $B_{35\%}$ but below $B_{35\%}$, the determination depends on the mean spawning biomass for `r year+12` 
If the mean spawning biomass for `r year+12` is below $B_{35\%}$, the stock is approaching an overfished condition. 
Otherwise, the stock is not approaching an overfished condition.
Based on the above criteria and Table 10.16, the stock is not overfished and is not approaching an overfished condition.


Based on the above criteria and Table 12-16, the stock is not overfished and is not approaching an overfished condition. 
The test for determining whether a stock is overfished is based on the 2019 catch compared to OFL. 
The official total catch for 2019 is 2,491 t which is less than the 2019 OFL of 4,521 t; therefore, the stock is not being subjected to overfishing. 
The tests for evaluating whether a stock is overfished or approaching a condition of being overfished require examining model projections of spawning biomass relative to B35% for 2020 and 2022. 
The estimates of spawning biomass for 2020 and 2022 from the current year (2020) projection model are 37,587 t and 37,526 t, respectively. 
Both estimates are above the B35% estimate of 21,299 t and, therefore, the stock is not currently overfished nor approaching an overfished condition. 
The F that would have produced a catch for 2019 equal to the OFL of 2019 was 0.06.


The fishing mortality that would have produced a catch for last year equal to last year’s OFL is `r best_f`.


